\documentclass[./main.tex]{subfiles}
\begin{document}

One of the easiest way to
create new spaces from given spaces is
by taking ``subspaces''.
The first of these kind of spaces
we will consider is \emph{closed embedding}.
\footnote{
  This is more standardly called \emph{closed immersion}.
  I do not like the terminology because
  for those who have touched manifolds,
  the word ``immersion'' brings too much preconceptions.
}
The idea is simple : 
given $f \in A$ a $k$-algebra,
we want to consider the ``subspace of $\SPEC A$ where $f = 0$''.

\begin{dfn}
  
  Let $A$ be a $k$-algebra.
  Given an element $f \in A$,
  we define the \emph{zero locus of $f$} to be
  the affine scheme $V(f) := \SPEC A / (f)$.
  The quotient map $A \to A / (f)$
  corresponds to a morphism of affine scheme over $k$ \[
    i : \SPEC A / (f) \to \SPEC A
  \]

  More generally,
  given a morphism $i : Z \to X$ of affine schemes over $k$,
  we call it a \emph{closed embedding} when
  the corresponding algebra map $\cO(X) \to \cO(Z)$ is surjective.
  For such a morphism,
  the kernel $I$ of $\cO(X) \to \cO(Z)$
  is called \emph{the ideal of $i$}.
  We refer to $Z$ as the \emph{zero locus of $I$}.

  A \emph{closed subspace} of $X \in \AFF_k$
  is an isomorphism class of closed embeddings into $X$.

\end{dfn}

\begin{prop}
  
  Let $S \in \AFF_k$.
  Then we have an order reversing
  bijection between
  the set of closed embeddings into $S$ and
  the set of ideals of $\cO(S)$
  given by taking zero locus and taking ideals.
\end{prop}
\begin{proof}
  A tautology.
\end{proof}

\begin{prop}[Universal property of zero locus]
  
  Let $i : Z \to X$ be a closed embedding in $\AFF_k$
  and let $I$ be its ideal.
  Then for any morphism $\varphi : T \to S$ in $\AFF_k$,
  there is a factoring \begin{cd}
    T & Z \\
    & S 
    \arrow[from = 1-1, to = 1-2 , dashed]
    \arrow[from = 1-2 , to = 2-2 , "{i}"]
    \arrow[from = 1-1 , to = 2-2 , "{\varphi}"']
  \end{cd}
  if and only if under the $k$-algebra map
  $\varphi^* : \cO(S) \to \cO(T)$,
  the ideal $I$ is sent to zero.
  Furthermore,
  such a factoring is unique when it exists.
\end{prop}
\begin{proof}
  After dualizing affine schemes over $k$
  to $k$-algebras,
  this is precisely the universal property
  of the quotient map $\cO(S) \to \cO(S) / I$.
\end{proof}

In what sense is a closed embedding $i : Z \to X$ 
the ``injection of a subspace''?
For this we review subsets of sets.

\begin{prop}[Characterisation of monomorphisms in sets]
  
  Let $i : Z \to X$ be a map of sets.
  Then $i$ is an isomorphism onto a subset of $X$ iff
  for all maps $x , x_1 : T \to Z$ 
  such that $i(x) = i(x_1)$,
  we have $x = x_1$.
\end{prop}
\begin{proof}
  Exercise.
\end{proof}

The categorical property which generalises the above phenomenon
is the notion of a \emph{monomorphism}.
In fact, part of the
statement of the universal property of zero locus
is that the closed embedding is a 
monomorphism in $\AFF_k$.

\end{document}